### 2007-01-18 Documentation/part2.tex
JK - Minor modifications in the appendices (except appendix B)
*** empty log message ***

The description of the neutrino oscillation physics is, at least numerically, relatively simple. We use the {\em evolution operator method} (see, \eg, \Ref~\cite{Ohlsson:1999um})  to compute the neutrino oscillation probabilities and divide the matter density profile into layers of constant matter density. For each of these layers, the Hamiltonian in matter is diagonalized in order to propagate the neutrino transition amplitudes. Finally, the transition probability is obtained as the absolute square of the total neutrino transition amplitudes. Depending on the precision of the studied experiment, this approach turns out to be precise enough in Earth matter even if only a small number of matter density steps is used. Since we allow an uncertainty of the matter density profile, it is, in fact, in most cases sufficient to consider only one density step with the average matter density together with a matter density uncertainty~\cite{Ohlsson:2003ip}. Note that this approach may not be applicable to quickly varying extraterrestrial matter density profiles.

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|cllc|}
\hline
{\tt @builtin} & Description & Parameters & Min. version \\
\hline
1 & Neutrino factory $\mu^+$ decay & {\tt @parent\_energy} [GeV], & 2.0 \\
& & {\tt @stored\_muons} &  \\
2 & Neutrino factory $\mu^-$ decay & {\tt @parent\_energy} [GeV], & 2.0 \\
& & {\tt @stored\_muons} &  \\
3 & Beta beam, inv. beta decay & {\tt @end\_point} [GeV], & 3.0 \\
& & {\tt @stored\_ions}, &  \\
& & {\tt @gamma} & \\
4 & Beta beam, beta decay & {\tt @end\_point} [GeV], & 3.0 \\
& & {\tt @stored\_ions}, &  \\
& & {\tt @gamma} & \\
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:binfluxes} Built-in fluxes currently supported by \GLOBES . For details on
the beta beams, see \Ref~\cite{Huber:2005jk}. MR ABOUT BETA BEAMS: DOES THIS REALLY WORK FOR ALL ISOTOPES? IS THE DIFFERENCE BETWEEN 3 AND 4 CORRECTLY DESCRIBED? WHY IS THERE A NORMALIZATION FACTOR 50.9 IN THE FILES?}
\end{table}

For a built-in neutrino source, one has to specify which
built-in spectrum should be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
The currently available fluxes are described in \Tab~\ref{tab:binfluxes}.
For example, two built-in neutrino factory fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is 
\index{aedl}{nuflux@{\tt nuflux}!builtin@{\tt "@builtin}}
\index{aedl}{nuflux@{\tt nuflux}!parent energy@{\tt "@parent\_energy}}
\index{aedl}{nuflux@{\tt nuflux}!stored muons@{\tt "@stored\_muons}}
\index{aedl}{nuflux@{\tt nuflux}!end point@{\tt "@end\_point}}
\index{aedl}{nuflux@{\tt nuflux}!stored ions@{\tt "@stored\_ions}}
\index{aedl}{nuflux@{\tt nuflux}!gamma@{\tt "@gamma}}
\begin{quote}
{\tt nuflux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
MR: SOME MORE ELABORATION AND AN EXAMPLE FOR THE BETA BEAMS MAY BE APPROPRIATE HERE [WW].
%
\index{aedl}{nuflux@{\tt nuflux}!power@{\tt "@power}}
\index{aedl}{nuflux@{\tt nuflux}!norm@{\tt "@norm}}
\index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}
For a user-defined flux, one has to specify the file name:
\begin{quote}
{\tt nuflux(\#user)<}\\
{\tt \tb @flux\_file = "user\_file\_1.dat"\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
For more details on the flux definition, see page~\pageref{app:flux} in the appendix.

>>>>

The description of the neutrino oscillation physics is, at least numerically, relatively simple. We use the {\em evolution operator method} (see, \eg, \Ref~\cite{Ohlsson:1999um})  to compute the neutrino oscillation probabilities and divide the matter density profile into layers of constant matter density. For each of these layers, the Hamiltonian in matter is diagonalized in order to propagate the neutrino transition amplitudes. Since this step is computationally expensive, a specialized algorithm is used~\cite{Kopp:2006wp}. Finally, the transition probability is obtained as the absolute square of the total neutrino transition amplitudes. Depending on the precision of the studied experiment, this approach turns out to be precise enough in Earth matter even if only a small number of matter density steps is used. Since we allow an uncertainty of the matter density profile, it is, in fact, in most cases sufficient to consider only one density step with the average matter density together with a matter density uncertainty~\cite{Ohlsson:2003ip}. Note that this approach may not be applicable to quickly varying extraterrestrial matter density profiles.

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|cllc|}
\hline
{\tt @builtin} & Description & Parameters & Min. version \\
\hline
1 & Neutrino factory $\mu^+$ decay & {\tt @parent\_energy} [GeV], & 2.0 \\
& & {\tt @stored\_muons} &  \\
2 & Neutrino factory $\mu^-$ decay & {\tt @parent\_energy} [GeV], & 2.0 \\
& & {\tt @stored\_muons} &  \\
3 & Beta beam, inv. beta decay & {\tt @end\_point} [GeV], & 3.0 \\
& & {\tt @stored\_ions}, &  \\
& & {\tt @gamma} & \\
4 & Beta beam, beta decay & {\tt @end\_point} [GeV], & 3.0 \\
& & {\tt @stored\_ions}, &  \\
& & {\tt @gamma} & \\
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:binfluxes} Built-in fluxes currently supported by \GLOBES . For details on
the beta beams, see \Ref~\cite{Huber:2005jk}. }
\end{table}

For a built-in neutrino source, one has to specify which
built-in spectrum should be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
The currently available fluxes are described in \Tab~\ref{tab:binfluxes}.
For example, two built-in neutrino factory fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is 
\index{aedl}{nuflux@{\tt nuflux}!builtin@{\tt "@builtin}}
\index{aedl}{nuflux@{\tt nuflux}!parent energy@{\tt "@parent\_energy}}
\index{aedl}{nuflux@{\tt nuflux}!stored muons@{\tt "@stored\_muons}}
\index{aedl}{nuflux@{\tt nuflux}!end point@{\tt "@end\_point}}
\index{aedl}{nuflux@{\tt nuflux}!stored ions@{\tt "@stored\_ions}}
\index{aedl}{nuflux@{\tt nuflux}!gamma@{\tt "@gamma}}
\begin{quote}
{\tt nuflux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
%
Furthermore, two beta beam neutrino fluxes are available: inverse beta decay ({\tt @builtin = 3}, \ie\ from the decay of
$^{18}$Ne-isotopes) and beta decay ({\tt @builtin = 4}, \ie\ from the decay of $^{6}$He-isotopes). For a beta beam file, using
the beta beam built-in fluxes, the end point energy of the decay ({\tt @end\_point}), the number of decays per year 
({\tt @stored\_muons}), and the acceleration factor of the ions $\gamma$ ({\tt @gamma}) have to be specified. An example for
setting up a beta beam neutrino flux from the decay of $^{18}$Ne isotopes is
\begin{quote}
{\tt nuflux(\#nu\_e\_flux)<\\
\tb  @builtin = 3\\
\tb  @gamma = 130.0\\
\tb  @end\_point = 0.0034\\
\tb  @stored\_muons = 2.2e+18\\
\tb  @time = 4.0\\
>}
\end{quote}
%
\index{aedl}{nuflux@{\tt nuflux}!power@{\tt "@power}}
\index{aedl}{nuflux@{\tt nuflux}!norm@{\tt "@norm}}
\index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}
%
For a user-defined flux, one has to specify the file name:
\begin{quote}
{\tt nuflux(\#user)<}\\
{\tt \tb @flux\_file = "user\_file\_1.dat"\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
For more details on the flux definition, see page~\pageref{app:flux} in the appendix.
